HGP analysis: fit model

### 2019-03-26 PlottingScripts/ProcessPlotData.Rmd
update

m <- merge(Reg,
           NumPop, 
           by = c('Chr','Pos'))
#count number of significant sites per population
n <- as.data.frame(table(m$Pop))
#put it all together for plotting
overLap <- unique(merge(m,n,
                        by.x = 'Population',
                        by.y = 'Var1'))
#include number of significant sites in titles
overLap$title = paste0(overLap$Pop,
                       ' : ', 
                       overLap$Freq)
#add asterix for Phase 1 pops
Phase1Title <- paste('*',
                     overLap[which(overLap$Pop 
                                   %in% phase1),]$title)
overLap[which(overLap$Pop
              %in% phase1),]$title <- Phase1Title
#set order by SigHits, Chromosome and Position
overLap <- overLap[with(overLap, 
                        order(-SigHits,
                              Chr,
                              Pos)),]
overLap$Pos <- factor(overLap$Pos, 
                      levels = unique(overLap$Pos))
#this is for keeping track of the number of variants
overLap$nrow <- seq(length = nrow(overLap))
#this is for discrete point size
overLap$Pcat <- 4
overLap[which(overLap$plog10 < 8),]$Pcat <- 2
overLap[which((overLap$plog10 > 8) & 
              (overLap$plog10 < 10)),]$Pcat <- 3
#this is for the labels of the xaxis
xaxis <- overLap[!duplicated(overLap[,c('SigHits')]),]
xaxis <- xaxis[which(xaxis$SigHits < 11),]
xaxis <- rbind(xaxis,tail(overLap, 1))

>>>>

m <- merge(Reg,
           NumPop, 
           by = c('Chr','Pos'))
#count number of significant sites per population
n <- as.data.frame(table(m$Pop))
#put it all together for plotting
overLap <- unique(merge(m,n,
                        by.x = 'Population',
                        by.y = 'Var1'))
#include number of significant sites in titles
overLap$title = paste0(overLap$Pop,
                       ' : ', 
                       overLap$Freq)
#add asterix for Phase 1 pops
Phase1Title <- paste('*',
                     overLap[which(overLap$Pop 
                                   %in% phase1),]$title)
overLap[which(overLap$Pop
              %in% phase1),]$title <- Phase1Title
#set order by SigHits, Chromosome and Position
overLap <- overLap[with(overLap, 
                        order(-SigHits,
                              Chr,
                              Pos)),]
overLap$Pos <- factor(overLap$Pos, 
                      levels = unique(overLap$Pos))
#this is for keeping track of the number of variants
overLap$nrow <- seq(length = nrow(overLap))
#this is for discrete point size
overLap$Pcat <- 4
overLap[which(overLap$plog10 < 8),]$Pcat <- 2
overLap[which((overLap$plog10 > 8) & 
              (overLap$plog10 < 10)),]$Pcat <- 3

### 2019-03-26 rick/final/3_fit_model_compareModels.Rmd
update

#plinkPath <- args[2] 
plinkPath <- '/Users/luke/genomes/genomes/hg19/plink/'

#chr <- args[3]
chr <- '22'

@@

```{r}
library(tidyverse)
library(lfa)
library(furrr)
library(glm2)
library(tictoc)
library(data.table)
library(cowplot)
options(future.globals.maxSize= 1000000000000)
plan(multicore)
```
load plink format genotype data
```{r}
data_genotype <- read.bed(paste0(plinkPath,'chr',chr))

@@

rownames(data_genotype) <- data_bim$rsID
```
run lfa if needed
```{r lfa}
tic()
#LF <- lfa(data_genotype, 5)
toc()
message("lfa run complete")
```
save lfa if needed
```{r save}
tic()
#saveRDS(LF, "./data/preprocessed/lf5.rds")
toc()
```
load logistic factor analysis results
```{r read lf}
LF <- readRDS( "./data/preprocessed/lf5.rds")
```

@@

```

```{r}

@@

write.table(data_fit[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_LFA.txt")),
              quote = FALSE,
              row.names = FALSE)
data_fit[,2]<-NULL
```
#PCA
```{r, warning=FALSE}
#load PCA
library(data.table)
samples = fread('/Users/luke/Documents/PCAperPop/Name_Pop_Qual_PC1_PC2_PC3_PC4_PC5.txt')
PC <- as.matrix(samples[,c('PC1','PC2','PC3','PC4','PC5')])
PC <- cbind(PC, 1)

@@

data_fit_PCA_dich <- 
	data_genotype %>%
	as_tibble() %>%
	transmute(
	fit = future_pmap(., fit_glm_dich,
	LF= PC, 
	trait=avg_qual_mean),
		rsID=labels)  %>%
	unnest(fit) %>%
	mutate(l= rep(c("data","stat"), n()/2))%>%
	spread(l, fit)%>%
	unnest(stat)
   toc()    
```

```{r}

@@

write.table(data_fit_PCA[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_PCA.txt")),
              quote = FALSE,
              row.names = FALSE)
data_fit_PCA[,2]<-NULL
```
#Population
```{r, warning=FALSE}

@@

```
```{r}

@@

data_fit_Pop[,2]<-NULL
```

@@

```{r,fig.height = 10, fig.width = 10}

plt <- merge(data_fit[-2], 
             data_fit_PCA[-2], 
             by=c('rsID'), 
             suffixes = c('LFA','PCA'))
plt <- merge(plt, 
             data_fit_Pop[-2], 
             by=c('rsID'), 
             suffixes = c('','Pop'))

beta1<-ggplot(plt, aes(b, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Beta')
beta2<-ggplot(plt, aes(bLFA, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Beta')
beta3<-ggplot(plt, aes(bLFA, b))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Beta')
betas <- plot_grid(beta1,beta2,beta3,nrow=1)

af1<-ggplot(plt, aes(1-num_zeros/2504, bLFA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='AF', y='LFA')+ggtitle('Beta')+xlim(c(0,0.3))
af2<-ggplot(plt, aes(1-num_zeros/2504, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='AF', y='PCA')+ggtitle('Beta')+xlim(c(0,0.3))
af3<-ggplot(plt, aes(1-num_zeros/2504, b))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='AF', y='Pop')+ggtitle('Beta')+xlim(c(0,0.3))
af <- plot_grid(af1,af2,af3,nrow=1)
af
ggsave('./Beta_AF.jpg', height=4, width=8)

plt <- plt[which(plt$num_zeros < 2501),]

beta1<-ggplot(plt, aes(b, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Beta')
beta2<-ggplot(plt, aes(bLFA, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Beta')
beta3<-ggplot(plt, aes(bLFA, b))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Beta')
betas2 <- plot_grid(beta1,beta2,beta3,nrow=1)

af1<-ggplot(plt, aes(1-num_zeros/2504, bLFA))+geom_point(shape=1)+labs(x='AF', y='LFA')+ggtitle('Beta')+xlim(c(0,0.3))
af2<-ggplot(plt, aes(1-num_zeros/2504, bPCA))+geom_point(shape=1)+labs(x='AF', y='PCA')+ggtitle('Beta')+xlim(c(0,0.3))
af3<-ggplot(plt, aes(1-num_zeros/2504, b))+geom_point(shape=1)+labs(x='AF', y='Pop')+ggtitle('Beta')+xlim(c(0,0.3))
af2 <- plot_grid(af1,af2,af3,nrow=1)
af2
ggsave('./Beta_AF2.jpg', height=4, width=8)

makePlot <- function(x,y,xlab,ylab,title) {
p <- ggplot(plt, aes(x, y)) + 
  geom_abline(color='grey') +
  geom_point(shape=1) +
  labs(x=xlab, y=ylab) +
  ggtitle(title)

return(p)
}

p1 <- makePlot(plt$dev0LFA, 
               plt$dev0PCA, 
               'Null Deviance LFA', 
               'Null Deviance PCA', 
               'Null Model')
p2 <- makePlot(plt$devdiffLFA, 
               plt$devdiffPCA, 
               'Deviance LFA', 
               'Deviance PCA', 
               'Deviance')

p2.2<-ggplot(plt, aes(devdiffLFA, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))

p3<-ggplot(plt, aes(bLFA, bPCA, color=1-num_zeros/2504))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Beta')
pp1 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(plt, aes(dev0LFA, dev0))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiffLFA, devdiff))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiffLFA, devdiff))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(plt, aes(bLFA, b, color=1-num_zeros/2504))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Beta')
pp2 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(plt, aes(dev0, dev0PCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiff, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiff, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Deviance Zoom')+xlim(c(0,50))+ylim(c(0,50))
p3<-ggplot(plt, aes(b, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Beta')
pp3 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

plot_grid(pp1,pp2,pp3,ncol=3)
ggsave('./CompareModels.jpg', height=10, width = 10)

```

```{r,fig.height = 10, fig.width = 10}

plt <- merge(data_fit_dich[-2], data_fit_PCA_dich[-2], by=c('rsID'), suffixes = c('LFA','PCA'))
plt <- merge(plt, data_fit_Pop_dich[-2], by=c('rsID'), suffixes = c('','Pop'))

p1<-ggplot(plt, aes(dev0LFA, dev0PCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiffLFA, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiffLFA, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(plt, aes(bLFA, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Beta')
pp1 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(plt, aes(dev0LFA, dev0))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiffLFA, devdiff))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiffLFA, devdiff))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(plt, aes(bLFA, b))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Beta')
pp2 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(plt, aes(dev0, dev0PCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiff, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiff, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Deviance Zoom')+xlim(c(0,50))+ylim(c(0,50))
p3<-ggplot(plt, aes(b, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Beta')
pp3 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

plot_grid(pp1,pp2,pp3,ncol=3)
ggsave('./CompareModels_dich.jpg', height=10, width = 10)

```

```{r,fig.height = 10, fig.width = 10}

LFAp <- merge(data_fit[-2], data_fit_dich[-2], by=c('rsID'), suffixes = c('Poly','Dich'))
PCAp <- merge(data_fit_PCA[-2], data_fit_PCA_dich[-2], by=c('rsID'), suffixes = c('Poly','Dich'))
Popp <- merge(data_fit_Pop[-2], data_fit_Pop_dich[-2], by=c('rsID'), suffixes = c('Poly','Dich'))

p1<-ggplot(LFAp, aes(dev0Poly, dev0Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Null Model LFA')
p2<-ggplot(LFAp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance')
p2.2<-ggplot(LFAp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(LFAp, aes(bPoly, bDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Beta')
pp1 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(PCAp, aes(dev0Poly, dev0Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Null Model PCA')
p2<-ggplot(PCAp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance')
p2.2<-ggplot(PCAp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(PCAp, aes(bPoly, bDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Beta')
pp2 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(Popp, aes(dev0Poly, dev0Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Null Model Pop')
p2<-ggplot(Popp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance')
p2.2<-ggplot(Popp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(Popp, aes(bPoly, bDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Beta')
pp3 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

plot_grid(pp1,pp2,pp3,ncol=3)
ggsave('./CompareModels_VSdich.jpg', height=10, width = 10)

```

```{r,fig.height = 10, fig.width = 6}
#plot diff
plt <- merge(data_fit[-2], data_fit_Pop_dich[-2], by=c('rsID'), suffixes = c('LFA_Full','Pop_Dich'))

p1<-ggplot(plt, aes(dev0LFA_Full, dev0Pop_Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA Full', y='Pop Dich')+ggtitle('Null Model')+coord_flip()
p2<-ggplot(plt, aes(devdiffLFA_Full, devdiffPop_Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA_Full', y='Pop_Dich')+ggtitle('Deviance')+coord_flip()
p2.2<-ggplot(plt, aes(devdiffLFA_Full, devdiffPop_Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA_Full', y='Pop_Dich')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))+coord_flip()
p3<-ggplot(plt, aes(bLFA_Full, bPop_Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA_Full', y='Pop_Dich')+ggtitle('Beta')+coord_flip()
pp1 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

pp1
ggsave('./CompareModels_FullLFA_Popdich.jpg', height=10, width = 6)
```

>>>>

#plinkPath <- args[2] 
plinkPath <- '/Users/luke/Documents/QualityPaper/ProcessedData' #'/Users/luke/genomes/genomes/hg19/plink/'

#chr <- args[3]
chr <- 'PublishedVariants'

@@

```{r}
library(tidyverse)
library(lfa)
library(furrr)
library(glm2)
library(tictoc)
library(data.table)
library(cowplot)
options(future.globals.maxSize= 25000000000)
plan(multicore)
```
load plink format genotype data
```{r}
data_genotype <- read.bed(paste0(plinkPath,'chr',chr))

@@

data_bim <- 
  data_bim %>%  
  rowid_to_column() %>% 
  mutate(rsID= if_else(!duplicated(rsID),
                       rsID,
                       paste(rsID,
                             rowid,
                             sep="_" ))) %>% 
  select(-rowid)

rownames(data_genotype) <- data_bim$rsID

#data_genotype <- head(data_genotype, 50000)
```
run lfa if needed
```{r lfa}
tic()
#LF <- lfa(data_genotype, 5)
toc()
message("lfa run complete")
```
save lfa if needed
```{r save}
tic()
#saveRDS(LF, "./data/preprocessed/lf5.rds")
toc()
```
load logistic factor analysis results
```{r read lf}
LF <- readRDS( "./data/preprocessed/lf5.rds")
```

@@

@@

write.table(data_fit[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_LFA.txt")),
              quote = FALSE,
              row.names = FALSE)
rm(data_fit)
```
#PCA
```{r, warning=FALSE}
#load PCA
library(data.table)
samples = fread('/Users/luke/Documents/PCAperPop/Name_Pop_Qual_PC1_PC2_PC3_PC4_PC5.txt')
PC <- as.matrix(samples[,c('PC1','PC2','PC3','PC4','PC5')])
PC <- cbind(PC, 1)

@@

@@

write.table(data_fit_PCA[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_PCA.txt")),
              quote = FALSE,
              row.names = FALSE)
rm(data_fit_PCA)
```
#Population
```{r, warning=FALSE}

@@

@@

rm(data_fit_Pop)
```

@@
